---
title: "README"
author: "Garrick Aden-Buie"
date: "6/14/2017"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## ðŸ¿ sqrrrl

`sqrrrl` is a small collection of utility functions that help build text-based SQL queries in an R-style native-feeling and functional manner.

Unlike other packages that build SQL queries using an object-oriented style, `sqrrrl` provides small functions that produce SQL snippets and can be linked together to compose SQL queries.
The result is that the code to produce the SQL statement reads much like the SQL statement iteself.
On the other hand, `sqrrrl` doesn't know anything about your database and can't help you out with completions, etc.

## Installation

```r
# install.packages('devtools')
devtools::install_github('gadenbuie/sqrrrl')
```

## Quick Overview

```{r quick overview}
library(sqrrrl)

# ---- SELECT ----
SELECT()
SELECT('col1', 'col2', 'col3')
SELECT(newcol = 'col1', avg_col1 = 'mean(col1)')
SELECT(letters[1:3], 't2' = letters[4:6])
SELECT(a = 'apple', b = 'banana', c = 'cherry')
SELECT('t1' = c(a = 'apple', b = 'banana'), c = 'cherry')
SELECT('t1' = c(a = 'apple', b = 'banana'), c = 'cherry', 't2' = c(d = 'dragon_fruit'))

# ---- FROM ----
FROM('table1')
FROM('table1', 'table2')
FROM(t1 = 'table1', t2 = 'table2', 'table3')

# ---- WHERE ----
WHERE('col1 IS NOT NULL')
WHERE(cond = TRUE, 'col1 = 2', 'col2 >= 10')
WHERE(cond = FALSE, 'col1 = 2', 'col2 >= 10')

# ---- IN ----
'id' %IN% 1:4
'id' %IN% letters[1:3]
# Note: if left-hand-side length == 1, then LHS is unquoted (for subqueries)
'id' %IN% paste(SELECT('id'), FROM('other_table'))
'in' %IN% quotes(letters[1])

# ---- LIKE ----
'text_col' %LIKE% 'Prefix%'
'text_col' %LIKE% c('Prefix1%', 'Prefix2%')

# ---- Comparison Operators ----
eq(id = 4)
neq(id = 4)
lt(id = 4)
leq(id = 4)
gt(id = 4)
geq(id = 4)

# ---- Contatenation Operators ----
AND(eq(id = 3, class = 'text_value'), geq(date = '2017-06-14'))
OR(eq(id = 9, id = 12), leq(id = 5))
# Concatenate snippets with the %+% infix operator
SELECT() %+% FROM('table') %+% WHERE(eq(id = 10))

# ---- Complete Example ----
(example_query <- 
  SELECT('mpg', weight = 'wt', cylinders = 'cyl') %+%
  FROM('mtcars') %+%
  WHERE(
    lt(gear = 4),
    geq(mpg = 15),
    'cyl' %IN% c(4,6)
  ))
```

## Formatting SQL Queries

`sqrrrl` also provides a simple wrapper for <https://github.com/andialbrecht/sqlparse>, a Python package for formatting SQL queries.
`sqlparse` can be installed via `pip install --upgrade sqlparse`, thereafter making available the system command `sqlformat`.

`sqrrrl::sqlformat()` pretty-prints SQL queries, such as the one above.

```{r sqlformat, results='asis'}
example <- sqlformat(example_query, header = 'A Beautifully Formatted Example Query')
cat(example)
```


## More Operators and Examples

```{r sql-more}
# ---- Joins ----
# Using one ID
JOIN('left_tbl', 'right_tbl', 'id')
LEFT_JOIN('l', c('r' = 'right_tbl'), 'id')
RIGHT_JOIN('l', c('r' = 'right_tbl'), 'id', prefer_using = FALSE)
# Join on multiple columns, with different names on left and right
JOIN(type = 'natural right', 'left_tbl', 'right_tbl', c('left.col1' = 'right.col1', 'id2'))
# Join multiple tables on same column
INNER_JOIN('left_tbl', c('right_1', 'right_2'), 'id_col')
# Join multiple tables on different columns
OUTER_JOIN('l', c(r1 = 'right_1', r2 = 'right_2'), list('col1', 'col2'))
# Join multiple tables on different coluns with different column names
JOIN('l', c(r1 = 'right_1', r2 = 'right_2'), list(c(right_1_id = 'id', c(right_2_id = 'id'))))

# ---- INSERT INTO ... VALUES ----
iris_example <- iris[c(1, 51, 101), ]
# Insert all rows & columns from a data.frame
INSERT_INTO_VALUES('iris', iris_example)
# Insert select columns from a data.frame
INSERT_INTO_VALUES('iris', iris_example, c('Petal.Length', 'Petal.Width', 'Species'))
# Insert named vector
INSERT_INTO_VALUES('iris', c('Petal.Length' = 1.9, 'Petal.Width' = 0.2, 'Species' = 'setosa'))
# Insert subset of named vector
INSERT_INTO_VALUES('iris', c('Petal.Length' = 1.9, 'Petal.Width' = 0.2, 'Species' = 'setosa'),
                   cols = c('Petal.Width', 'Species'))
# Insert just vector of mixed type without column names
INSERT_INTO_VALUES('iris', c(6.5, 3.2, 5.1, 2, 'virginica'))
```
